\section{Alternating Tensors and Exterior Powers}

\begin{definition}[Alternating map]
\label{def: alternating map}
Let \(V\) and \(W\) be vector spaces. A multilinear map \(f: V^n \to W\) is
called alternating if for any \((v_1, \dots, v_n) \in V^n\) such that exists \(i
< j\) for which \(v_i = v_j\) then
\[
  f(v_1, \dots, v_n) = 0.
\]
\end{definition}

\begin{proposition}\label{prop: alternating map property}
Let \(f: V^n \to W\) be an alternating multilinear map. Then if \(\tau \in
S_n\) is a transposition, we have
\[
  f(v_{\tau(1)}, \dots, v_{\tau(n)}) = - f(v_1, \dots, v_n).
\]
In general, if \(\sigma \in S_n\) is any permutation, then
\[
  f(v_{\sigma(1)}, \dots, v_{\sigma(n)}) = \sign(\sigma) f(v_1,
  \dots, v_n).
\]
\end{proposition}

\begin{proof}
Let \(\tau\) be a transposition between indices \(i < j\), then consider
\begin{align*}
  0 = f(v_1, \dots, v_i + v_j, \dots, v_i + v_j, \dots, v_n)
  =\, & f(v_1, \dots, v_i, \dots, v_i + v_j, \dots, v_n) \\
    & + f(v_1, \dots, v_j, \dots, v_i + v_j, \dots, v_n)
\end{align*}
therefore we find
\begin{align*}
  f(v_1, \dots, v_i, \dots, v_i + v_j, \dots, v_n)
  &= - f(v_1, \dots, v_j, \dots, v_i + v_j, \dots, v_n) \\
  f(v_1, \dots, v_i, \dots, v_j, \dots, v_n)
  &= - f(v_1, \dots, v_j, \dots, v_i, \dots, v_n) \\
  f(v_1, \dots, v_n)
  &= f(v_{\tau(1)}, \dots, v_{\tau(n)})
\end{align*}
where we used that \(f(v_1, \dots, v_i, \dots, v_i, \dots, v_n) = f(v_1,
\dots, v_j, \dots, v_j, \dots, v_n) = 0\). This shows the first proposition.
For the second proposition, by means of \cref{prop: sign is a group
homomorphism} and \cref{prop: permutations to transpositions} and the above
proposition for transpositions, we conclude the proof.
\end{proof}

\begin{remark}
Notice that \cref{prop: alternating map property} is not sufficient to
characterize alternating maps, it is but a necessary property. A
counterexample to that is the field with characteristic \(2\).
\end{remark}

\subsection{Alternating Tensor and Exterior Powers}

\begin{definition}[Alternating tensor]
\label{def: alternating tensor}
Let \(V\) be a \(k\)-vector space we define a tensor \(T \in T_0^q(V)\) to be
alternating if for all \(\sigma \in S_q\) permutation we have
\[
  f_\sigma(T) = \sign(\sigma) T.
\]
We denote by \(\Lambda^q V\) the subspace of \(T_0^q(V)\) of alternating
tensors and call it the \(q\)-exterior power of \(V\).
\end{definition}

\begin{proposition}[Alternating tensor projection]
\label{prop: alternating projection}
Let \(V\) be a \(k\)-vector space and \(\Char k \nmid q!\). We
define the linear projection operator \(A: T_0^q(V) \epi T_0^q(V)\) where
\[
  A(T) = \frac{1}{q!} \sum_{\sigma \in S_q} \sign(\sigma)
  f_\sigma(T)
\]
Then \(A^2 = A\) and \(\im A = \Lambda^q V\).
\end{proposition}

\begin{proof}
First we show that \(\im A \subseteq \Lambda^q V\) (we already know that
clearly \(\Lambda^q V \subseteq \im A\)). Notice that
\begin{align*}
  f_\sigma(A(T))
  &= f_\sigma \left( \frac 1 {q!} \sum_{\tau \in S_q}
  \sign(\tau) f_\tau(T) \right) \\
  &= \frac 1 {q!} \sum_{\tau \in S_q} \sign(\tau)
  f_{\sigma \tau}(T) \\
  &= \sign(\sigma) \left( \frac 1 {q!} \sum_{\tau \in \mathcal
  S_q} \sign(\sigma \tau) f_{\sigma \tau}(T) \right) \\
  &= \sign(\sigma) A(T)
\end{align*}
where we've used \cref{prop: sign is a group homomorphism}. Hence we conclude
that \(\im A = \Lambda^q V\). For the second proposition, notice that
\[
  A^2 = \frac 1 {q!^2} \sum_{\sigma, \tau \in S_q}
  \sign(\sigma \tau) f_{\sigma \tau}
  = \frac 1 {q!} \sum_{\rho \in S_q} \sign(\rho)
  f_\rho = A
\]
since any permutation \(\rho \in S_q\) can be represented in \(q!\)
different ways as a form of a product \(\sigma \tau\). This concludes the
proof.
\end{proof}

\begin{definition}[Exterior multiplication]
\label{def: exterior multiplication}
Let \(V\) be a \(k\)-vector space and \(v_1 \otimes \dots \otimes v_q \in
T_0^q(V)\). We define the exterior multiplication to be the map
\[
  A(v_1 \otimes \dots \otimes v_q) \coloneqq v_1 \wedge \dots \wedge v_q
\]
\end{definition}

\begin{proposition}[Universal property for exterior power]
\label{prop: exterior power universal property}
Let \(V\) be a \(k\)-vector space. For all \(k\)-vector spaces \(L\) together
with an alternating multilinear map \(\mu: V^q \to L\), there exists a unique
\(k\)-linear map \(\ell: \Lambda^q V \to L\) for which the diagram commutes
\[
  \begin{tikzcd}
    V^q \ar[d, swap, "\otimes"] \ar[r, "\mu"]  &L \\
    T_0^q(V) \ar[r, swap, two heads, "A"]
    &\Lambda^q V \ar[u, dashed, swap, "\ell"]
  \end{tikzcd}
\]
\end{proposition}

\begin{proof}
Since \(\Lambda^q V\) is a subspace of \(T_0^q(V)\), then we can use the
universal property \cref{thm: universal property of tensor products}.
\end{proof}

Now we get some geometric motivation behind the algebraic structure of the
exterior power \(\Lambda^q V\). Alternating tensors \(v_1 \wedge \dots \wedge
v_q \in \Lambda^q V\) can be seen as \(q\)-dimensional oriented volume elements,
where by oriented we mean that the transposition of two edges, say \(v_j\) and
\(v_i\), implies in a change up to a minus sign of the value (the sign is what
creates the bridge between orientation and wedge product). Notice that when
there are equal edges, the volume becomes malformed and \(v_1 \wedge \dots
\wedge v_q = 0\).

\begin{proposition}
Let \(V\) be a \(n\)-dimensional \(k\)-vector space, where
\(\operatorname{char} k \neq 2\). Define \(\{e_i\}_{i=1}^n\) to be a base for
\(V\). Then the exterior product \(e_{i_1} \wedge \dots \wedge e_{i_q} = 0\)
if there exists \(i_a = i_b\) for some \(1 \leq a, b \leq q\).
\end{proposition}

\begin{proof}
Since \(A\) is an alternating map, we can use \cref{def: alternating map} and
conclude the proof.
\end{proof}

For the next proposition, we proceed in a similar fashion as in \cref{prop:
basis for symmetric power}.

\begin{proposition}[Exterior power \(\Lambda^q V\) basis]
\label{prop: exterior power basis}
Let \(\{e_i\}_{i=1}^n\) be a basis for the \(k\)-vector space \(V\). The
factorizable tensors
\[
  A(e_{i_1} \otimes \dots \otimes e_{i_q}) = e_{i_1} \wedge \dots \wedge
  e_{i_q}
\]
with \(1 \leq i_1 < \dots < i_q \leq n\) form a basis for the subspace
\(\Lambda^q V\).
\end{proposition}

\begin{proof}
Let \(B \coloneq \{e_{i_1} \wedge \dots \wedge e_{i_q} \colon 1 \leq i_1 < \dots < i_q
\leq n\}\). Since \(\{e_{i_1} \otimes \dots \otimes e_{i_q} \colon 1 \leq i_1 <
\dots i_q \leq n\}\) generates \(T_0^q(V)\) (see \cref{lem: tensor basis}),
then clearly \(B\) does generate the subspace \(\Lambda^q V\). We now show
that \(B\) is linearly independent.

Denote by \(\mathcal I \coloneq \{I = (i_j)_{j=1}^q \colon 1 \leq i_1 < \dots < i_q \leq
n\}\) the set of strictly increasing \(q\)-tuples. For each \(I \coloneq (i_1,
\dots, i_q) \in \mathcal I\) we define the alternating multilinear map \(\mu_I
: V^q \to k\) as
\[
  \mu_I(v_1, \dots, v_q) = \sum_{\sigma \in S_q}
  \sign(\sigma) \prod_{j=1}^q e_{i_{\sigma(j)}}^*(v_{i_j})
\]
where \(\{e_{i_j}^*\}_{i_j \in I}\) is the dual of \(\{e_{i_j}\}_{i_j
\in I}\). Using \cref{prop: exterior power universal property} we can
define a unique linear map \(f_I: \Lambda^q V \to k\) such that the diagram
commutes
\[
  \begin{tikzcd}
    V^q \ar[r, "\mu_I"] \ar[d, swap, "A \circ \otimes"] &k \\
    \Lambda^q V \ar[ur, dashed, swap, "f_I"]
  \end{tikzcd}
\]
which implies in \(\mu_I(v_1, \dots, v_q) = f_I(v_1 \wedge \dots \wedge v_q)\)
for all \((v_1, \dots, v_q) \in V^q\).

Let \(I' = (i_1', \dots, i_q') \in \mathcal I\) such that \(I' \neq I\). From
the strictly ordering of the indices, we conclude that there must exists some
\(1 \leq j_0 \leq q\) such that \(i_{j_0}' \neq i_j\) for all \(1 \leq j \leq
q\). In particular, this implies that
\[
  f_I(e_{i_1'} \wedge \dots \wedge e_{i_q'}) = \mu_I(e_{i_1'}, \dots,
  e_{i_q'}) = \sum_{\sigma \in S_q} \sign(\sigma)
  \prod_{j=1}^q e_{i_{\sigma(j)}}^*(e_{i_j'}) = 0
\]
since \(e_{i_{\sigma(j)}}^*(e_{i_{j_0}}) = 0\) for all permutations \(\sigma
\in S_q\) and \(1 \leq j \leq q\). On the other hand, we have that
\[
  f_I(e_{i_1} \wedge \dots \wedge e_{i_q}) = \mu_I(e_{i_1}, \dots, e_{i_q})
  = \sum_{\sigma \in S_q} \sign(\sigma) \prod_{j=1}^q
  e_{i_{\sigma(j)}}^*(e_{i_j})
  = \sign(\Id) \prod_{j=1}^q e_{i_{\Id(j)}}^*(e_{i_j})
  = 1
\]

Let \(c_P \in k\) for all \(P \in \mathcal I\) and consider the vanishing
linear combination
\begin{equation}\label{eq: vanishing exterior product combination}
  \sum_{P \in \mathcal I} c_P (e_{p_1} \wedge \dots \wedge e_{p_q}) = 0.
\end{equation}
Then, if we look at its image under the map \(f_I\) for all \(I \in \mathcal
I\) we conclude that
\[
  0 = f_I\left( \sum_{P \in \mathcal I} c_P (e_{p_1} \wedge \dots \wedge
  e_{p_q}) \right) = \sum_{P \in \mathcal I} c_P f_I(e_{p_1} \wedge \dots \wedge
  e_{p_q}) = c_I
\]
hence we conclude that the linear combination \cref{eq: vanishing exterior
product combination} vanishes if and only if each coefficient vanishes. This
concludes that \(B\) is linearly independent. Hence we've proved that \(B\) is
a basis for \(\Lambda^q V\).
\end{proof}

\begin{proposition}
Let \(V\) be a \(n\)-dimensional \(k\)-vector space. Then we have
\[
  \dim_k \Lambda^q V = \binom n q
\]
\end{proposition}

\begin{proof}
Since we want our indices to be strictly increasing, we are left with \(n\)
elements of which we want to arrange in groups of \(q\) elements. Hence the
number of possible arrangements is \(\binom n q\) and this concludes the
proof.
\end{proof}

\subsection{Exterior Algebra}

\begin{definition}[Exterior algebra]
\label{def: exterior algebra}
Let \(V\) be a \(k\)-vector space. We define the exterior algebra on \(V\) as
\[
  \Lambda^\bullet V = \bigoplus_{q=0}^\infty \Lambda^q V
\]
together with a multiplicative structure \(\wedge: \Lambda^d V \otimes
\Lambda^q V \to \Lambda^{d+q} V\) mapping
\[
  (v_1 \wedge \dots \wedge v_d) \otimes (w_1 \wedge \dots \wedge w_q) \xmapsto
  \wedge v_1 \wedge \dots \wedge v_d \wedge w_1 \wedge \dots \wedge w_q.
\]
We interpret \(\Lambda^0 V = k\).
\end{definition}

\begin{proposition}\label{prop:exterior-algebra-associative-skew-commutative}
The multiplicative structure of the tensor algebra \(\Lambda^\bullet V\) is
associative and skew-commutative, that is, for all \(\alpha \in \Lambda^d V\)
and \(\beta \in \Lambda^q V\) we have \(\alpha \wedge \beta = (-1)^{d q} \beta
\wedge \alpha\).
\end{proposition}

\begin{proof}
First, we prove that if \(T_1 \in T_0^q(V)\) and \(T_2 \in T_0^d(V)\) then
\begin{equation}\label{eq: asso goal}
  A(A(T_1) \otimes T_2) = A(T_1 \otimes A(T_2)) =
  A(T_1 \otimes T_2)
\end{equation} Notice that \[
  A(T_1) \otimes T_2 = \frac 1 {q!} \sum_{\sigma \in S_q}
  \sign(\sigma) f_\sigma(T_1) \otimes T_2
\]
therefore we find
\begin{equation}\label{eq: A(A(T1) T2)}
  A(A(T_1) \otimes T_2) = \frac 1 {q!} \sum_{\sigma \in S_q}
  \sign(\sigma) A(f_\sigma(T_1) \otimes T_2)
\end{equation}
Moreover, we can construct an injection of the symmetry groups \(S_q
\mono S_{q + d}\) via the mapping \(\sigma \mapsto \overline \sigma\)
where \(\overline \sigma(i) = \sigma(i)\) for all \(i \in \{1, \dots, q\}\)
and \(\overline \sigma(i) = i\) for all \(i \in \{q + 1, \dots, q + d\}\). In
particular, we find that \(f_\sigma(T_1) \otimes T_2 = f_{\overline \sigma}
(T_1 \otimes T_2)\) and clearly \(\sign(\overline \sigma) = \sign(\sigma)\).
This implies that
\begin{equation}\label{eq: ol sigma and A commute}
  A(f_\sigma(T_1) \otimes T_2) = A(f_{\overline \sigma}(T_1 \otimes T_2))
  = f_{\overline \sigma}(A(T_1 \otimes T_2))
  = \sign(\overline \sigma) A(T_1 \otimes T_2)
  = \sign(\sigma) A(T_1 \otimes T_2)
\end{equation}

If we substitute \cref{eq: ol sigma and A commute} in \cref{eq: A(A(T1) T2)}
we find
\[
  A(A(T_1) \otimes T_2) = \frac 1 {q!} \sum_{\sigma \in S_q}
  \sign^2(\sigma) A(T_1 \otimes T_2) = A(T_1 \otimes T_2).
\]
In the same manner we can show that \(A(T_1 \otimes A(T_2)) = A(T_1 \otimes
T_2)\). Hence \cref{eq: asso goal} holds.

Let \(\alpha \in \Lambda^d V, \beta \in \Lambda^q V, \gamma \in \Lambda^p V\).
Notice that from the construction of the product map \(\wedge\) and \cref{eq:
asso goal} we find
\[
  (\alpha \wedge \beta) \wedge \gamma
  = \wedge\left( \wedge(\alpha \otimes \beta) \otimes \gamma \right)
  = \wedge\left(\alpha \otimes \wedge(\beta \otimes \gamma) \right)
  = (\alpha \wedge \beta) \wedge \gamma
\]
which proves associativity of the tensor algebra.

We now prove skew-commutativity. Let \(\alpha \in \Lambda^d V\) and \(\beta
\in \Lambda^q V\), and a permutation \(\sigma \in S_{q + d}\) such
that \(\sigma(i) = q + d - i + 1\), consisting of \(d q\) transpositions.
Hence we find that
\[
  \beta \wedge \alpha = f_\sigma(\alpha \wedge \beta)
  = \sign(\sigma)(\alpha \wedge \beta) = (-1)^{d q} (\alpha \wedge \beta).
\]
\end{proof}

\begin{proposition}\label{prop: li iff wedge nonzero}
Let \(V\) be a \(k\)-vector space and \(\{v_i\}_{i=1}^d \subseteq V\). The
vectors \(v_1, \dots, v_d\) are linearly independent if and only if
\[
  \Lambda^d V \ni v_1 \wedge \dots \wedge v_d \neq 0.
\]
\end{proposition}

\begin{proof}
Let \(\{v_i\}_{i=1}^d\) be a set of linearly dependent vectors and choose
a set of not-all zero scalars \(\{c_i\}_{i=1}^d \subseteq k\) such that
\(\sum_{i=1}^d c_i v_i = 0\). Choose \(1 \leq j \leq d\) such that \(c_j
\neq 0\) and write \(v_j = -\sum_{i=1}^d \frac{c_i}{c_j} v_i\). Then we find
that (recall \(\mathcal M_0\) from our construction of the tensor product)
\begin{align}
  \nonumber
  v_1 \wedge \dots \wedge v_j \wedge \dots \wedge v_d
  &= v_1 \wedge \dots \wedge \left( - \sum_{i=1}^d \frac{c_i}{c_j} v_i \right)
  \wedge \dots \wedge v_d
  \\
  \nonumber
  &= \left( v_1, \dots, - \sum_{i=1}^d \frac{c_i}{c_j} v_i, \dots, v_d \right)
  + \mathcal M_0 \\
  &= -\sum_{i=1}^d \frac{c_i}{c_j} (v_1, \dots, v_i, \dots, v_d) + \mathcal
  M_0
  \\
  \label{eq: wedge repeated v_i}
  &= - \sum_{i=1}^d \frac{c_i}{c_j} (v_1 \wedge \dots \wedge v_i \wedge \dots
  v_d)
  \\
  \label{eq: wedge zero}
  &= 0
\end{align}
Where \cref{eq: wedge zero} comes from the fact that we have a
repeated \(v_i\) in the wedge product \cref{eq: wedge repeated v_i}.

Suppose \(\{v_i\}_{i=1}^d\) is a linearly independent set. Then from
\cref{prop: li to basis} we can build \(B \supseteq \{v_i\}_{i=0}^d\) such
that \(B\) is a basis for \(V\). From \cref{prop: exterior power basis}, we
find that \(\mathcal B = \{v_{i_1} \wedge \dots \wedge v_{i_d} \colon v_{i_j} \in B
\text{ and } i_1 < \dots < i_d\}\) is a basis for \(\Lambda^d V\). In
particular, \(v_1 \wedge \dots \wedge v_d \in \mathcal B\), hence necessarily
\(v_1 \wedge \dots \wedge v_d \neq 0\).
\end{proof}

Therefore one can trivially see that the kernel of the alternating projection
\(A\) is simply the collection of all decomposable tensors \(v_1 \otimes \dots
\otimes v_d\) such that the collection \(\{v_{1}, \dots, v_d\} \subseteq V\) is
linearly dependent.

\begin{corollary}\label{cor:kernel-alternating-projection}
The kernel of the alternating projection \(A: T^d_0(V) \epi \Lambda^d V\) is
given by the collection \(\{v_{1} \otimes \dots \otimes v_d \in T^d_0(V)\}\)
such that \(\{v_{1} \otimes \dots \otimes v_d\}\) is linearly dependent on
\(V\).
\end{corollary}

\begin{theorem}\label{thm:iso-measure-volume-ext}
Let \(V\) be a finite dimensional \(k\)-vector space. The map \(\phi: \Lambda^d
V^{*} \isoto (\Lambda^dV)^{*}\) given by the mapping
\[
f_1 \wedge \dots \wedge f_d \overset{\phi}\longmapsto \left(
  v_1 \wedge \dots \wedge v_d \mapsto
  \sum_{\sigma \in S_d} \sign(\sigma) \prod_{j=1}^d f_j(v_{\sigma(j)})
\right)
\]
is a \(k\)-linear isomorphism.
\end{theorem}

\begin{proof}
First of all, we show that \(\phi\) is a \(k\)-linear map. For each \(f \in
\Lambda^d V^{*}\), define the map \(\psi_{f} \in (\Lambda^dV)^{*}\) for which
\(\phi(f) = \psi_f\). Notice that if \(a \in k\) and \(f, g \in
\Lambda^d(V^{*})\), then \(\phi(f + a g) = \psi_{f + a g}\). Notice that
\begin{align*}
\psi_{f + ag}(v_1 \wedge \dots \wedge v_d)
&= \sum_{\sigma \in S_d} \sign(\sigma)
  \prod_{j=1}^d (f_{j} + a g_{j})(v_{\sigma(j)}) \\
&= \sum_{\sigma \in S_d} \sign(\sigma)
  \prod_{j=1}^d f_j(v_{\sigma(j)}) + a g_j(v_{\sigma(j)}) \\
&= \sum_{\sigma \in S_d} \sign(\sigma) \prod_{j=1}^d f_j(v_{\sigma(j)})
+ a \sum_{\sigma \in S_d} \sign(\sigma) \prod_{j=1}^d g_j(v_{\sigma(j)}) \\
&= \psi_f + a \psi_g
\end{align*}
Thus \(\phi(f + ag) = \phi(f) + a \phi(g)\) and therefore \(\phi\) is indeed
linear.

Suppose \(\dim V = n\) and let \(\{e_{j}\}_{j=1}^n\) be a base for \(V\). Define
the collection of strictly increasing \(d\)-tuples \(\mathcal I = \{I =
(i_{j})_{j=1}^d \colon 1 \leq i_1 < \dots < i_d \leq n\}\). From \cref{prop: exterior
power basis}, we know that \(\{e_{i_{1}} \wedge \dots \wedge e_{i_{d}} \in
\Lambda^d V \colon i_j \in I \text{ and } I \in \mathcal I\}\) is a basis for
\(\Lambda^d V\). Moreover, \(\{e_{j}^{*}\}_{j=1}^n\) is a basis for the dual
space \(V^{*}\), thus the collection \(\{e_{i_{1}}^{*} \wedge \dots \wedge
e_{i_{d}}^{*} \in \Lambda^d V \colon i_j \in I \text{ and } I \in \mathcal I\}\) is a
basis for \(\Lambda^d V^{*}\).

Let \(I \in \mathcal I\) be any strictly increasing \(d\)-tuple, and define the
notation \(e_I^{*} \coloneq e_{i_1}^{*} \wedge \dots \wedge e_{i_d}^{*}\), so
that
\begin{equation}
\label{eq:basislambda*}
e_I^{*} \overset{\phi}\longmapsto
\psi_{e_I^{*}}(v_1 \wedge \dots \wedge v_d)
= \sum_{\sigma \in S_{d}} \sign(\sigma) \prod_{j=1}^d e_{i_j}^{*}(v_{\sigma(j)}) =
\begin{cases}
  1, &\text{if }\ v_1 \wedge \dots \wedge v_d
        = e_{i_1} \wedge \dots \wedge e_{i_d} \\
  0, &\text{otherwise}
\end{cases}
\end{equation}
Notice that we mapped a basis of \(\Lambda^d V^{*}\) into a basis of
\((\Lambda^d V)^{*}\), since any map \(\psi \in (\Lambda^d V)^{*}\) can be
obtained by a unique linear combination of maps of the collection
\(\{\psi_{e_{I}^{*}} \in (\Lambda^d V)^{*} \colon I \in \mathcal I\}\) --- which,
together with \cref{eq:basislambda*}, makes it a basis. This shows that the
morphism \(\phi\) establishes an isomorphism \(\Lambda^dV^{*} \iso (\Lambda^d
V)\).
\end{proof}

\todo[inline]{
Add cool propositions for the exterior algebra: the Hodge Star Operator; the
isomorphisms
\begin{gather*}
\Sym^d(V \oplus W) \iso \bigoplus_{i=0}^d \Sym^i(V) \otimes \Sym^{d-i}(W) \\
\Lambda^d(V \oplus W) \iso
\bigoplus_{i=0}^d \Lambda^i(V) \otimes \Lambda^{d-i}(W)
\end{gather*}
}

\todo[inline]{After creating sections on affine and projective geometry,
introduce grassmanian varieties.}
