\section{ZFC}

\begin{theorem}[Cantor]
    \label{thm:cantor-power-set-theorem}
    Let \(A\) be a set and \(f: A \to 2^A\) be a map, then \(f\) is not surjective
    and hence \(|A| < |2^A|\).
\end{theorem}

\begin{proof}
    Let \(B \coloneq \{x \in A \colon x \notin f(x)\}\) be a set of \(2^A\) and
    suppose, for the sake of contradiction, that \(f\) is surjective. That is,
    there must exist some \(y \in A\) for which \(f(y) = B\) --- notice however that
    this can't be the case since by the law of excluded middle if \(y \in B\) then
    \(y \notin f(y)\) thus \(f(y) \neq B\), while if \(y \notin B\) then \(y \in
    f(y)\) and thus \(f(y) \neq B\). Therefore \(f\) cannot be surjective and thus
    \(|A| < |2^A|\) since there exists an injective mapping \(x \mapsto \{x\}\).
\end{proof}

\begin{theorem}
    \label{thm:no-set-contains-all-sets}
    There is no set containing all sets as members.
\end{theorem}

\begin{proof}
    For the sake of contradiction, let \(A\) be a set containing every set as a
    member. In particular \(2^A \subseteq A\) so that \(|2^A| \leq A\) --- this
    can't be the case by \cref{thm:cantor-power-set-theorem}, thus \(A\) cannot be a
    set.
\end{proof}

\begin{definition}[Cardinal numbers]
    \label{def:ordinal-cardinal}
    A \emph{cardinal number} is an isomorphism \emph{class} of sets, and the
    \emph{cardinality} of a given set \(S\) is its isomorphism class. The following
    are properties pertaining to cardinal numbers:
    \begin{enumerate}[(a)]\setlength\itemsep{0em}
        \item Every set has a unique cardinal number as its cardinality.
        \item Every cardinal number is the cardinality of some set.
        \item Two sets have the same cardinality if and only if they are isomorphic as
              sets.
    \end{enumerate}
\end{definition}

\begin{definition}[Well ordering]
    \label{def:well-ordering}
    A \emph{well-ordering} on a set \(S\) is a total ordering such that every
    non-empty subset of \(S\) has a \emph{least} element. A set equipped with a
    well-order is called a \emph{poset}.
\end{definition}

\begin{definition}[Ordinal]
    \label{def:ordinal}
    An \emph{ordinal number} is an isomorphism \emph{class} of \emph{well-ordered
        sets}, and the \emph{ordinal rank} of a poset \(S\) is its isomorphism
    class. The following are properties satisfied by the ordinal numbers in its
    regard to ordinal ranks:
    \begin{enumerate}[(a)]\setlength\itemsep{0em}
        \item Every poset has a unique ordinal numbers as its ordinal rank.

        \item Every ordinal number is the ordinal rank of some poset.

        \item Two sets have the same ordinal rank if and only if they are isomorphic as
              posets.
    \end{enumerate}
    Moreover, ordinals also have the following properties:
    \begin{enumerate}[(a)]\setcounter{enumi}{3}\setlength\itemsep{0em}
        \item Every ordinal \(\alpha\) has an immediate successor \(\alpha + 1\) ---
              this process entails the addition of an element to the end of a chain of a
              well-ordering of type \(\alpha\).

        \item There is a natural well-ordering on the collection of all ordinals. Given
              two ordinals \(\alpha\) and \(\beta\), we say that \(\alpha \leq \beta\) if
              and only if there exists an initial segment of \(\beta\) for which \(\alpha\)
              is isomorphic to.

        \item The induced well-ordering on the set \(\{\beta \colon \beta < \alpha\}\)
              is the isomorphism class represented by \(\alpha\) --- therefore, one can
              define an ordinal as a set containing all the smaller ordinals as members.

        \item (Equivalent to the Axiom of Choice) Every set is well-orderable, hence
              bijective to some ordinal.
    \end{enumerate}

    We denote by \(\omega\) the \emph{natural numbers}, so that every ordinal
    bijective to \(\omega\) is said to be \emph{countable}.
\end{definition}

The cardinal numbers are well-ordered, and can be indexed by means of
ordinals. We denote the \(\alpha\)-th cardinal number by \(\aleph_{\alpha}\) ---
hence \(\aleph_0 = \omega\), while \(\aleph_1 = \omega_1\) (the first
uncountable ordinal), for instance.

\begin{definition}[Successor \& limit ordinal]
    \label{def:successor-limit-ordinal}
    Given an ordinal \(\alpha\), we can classify it as a \emph{successor ordinal},
    if there exists an ordinal \(\beta\) such that \(\alpha = \beta + 1\), or as a
    \emph{limit ordinal} --- in particular, every cardinal is a limit ordinal.
\end{definition}

\begin{definition}[Successor \& limit cardinal]
    \label{def:successor-limit-cardinal}
    A cardinal is said to be a \emph{successor cardinal} if its corresponding
    indexing ordinal is a successor ordinal, otherwise it is a \emph{limit
        cardinal}.
\end{definition}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../deep-dive"
%%% End:
