\section{Torsion}

\begin{definition}[Torsion]
\label{def:torsion-module}
Let \(R\) be an integral domain, and \(M\) be an \(R\)-module. We define the
\emph{torsion} of \(M\) to be the submodule
\[
\torsion M \coloneq
\{m \in M \colon r m = 0 \text{ for some } r \in R \setminus 0\}.
\]
If \(\torsion M = 0\), we say that \(M\) is of \emph{free-torsion}. On the
contrary, if \(\torsion M = M\), then \(M\) is said to be of \emph{torsion}.
\end{definition}

\begin{definition}[Rank]
\label{def:rank-of-general-module}
Given an integral domain \(R\) and an \(R\)-module \(M\), we define the
\emph{rank} of \(M\) to be the cardinality of the maximal \(R\)-linearly
independent subset of \(M\).
\end{definition}

\begin{example}
\label{exp:torsion-module-has-rank-zero}
If \(R\) is an integral domain, and \(M\) is a torsion \(R\)-module, then
\(\rank_R M = 0\). Indeed, for any singleton \(\{m\} \subseteq M\) there exists
a non-zero \(r \in R\) such that \(r m = 0\), which shows that every singleton
is \(R\)-linearly dependent.
\end{example}

\section{Noetherian Rings \& Modules}

\subsection{Chain Conditions}

Recalling the definition of Noetherian rings (see \cref{def:noetherian-ring}),
we can extend this notion to the evironment of modules:

\begin{definition}
\label{def:noetherian-module}
Let \(R\) be a commutative ring. We say that an \(R\)-module \(M\) is
\emph{Noetherian} if every submodule of \(M\) is finitely generated.
\end{definition}

\begin{proposition}
\label{prop:equiv-conditions-noetherian}
Let \(R\) be a commutative ring, and \(M\) be an \(R\)-module. The following are
equivalent propositions:
\begin{enumerate}[(a)]\setlength\itemsep{0em}
\item \(M\) is a \emph{Noetherian} module.

\item Every \emph{ascending chain} of submodules of \(M\) \emph{stabilizes}. In
  other words, if \((N_j)_{j \in \N}\) is a collection of submodules of \(M\)
  such that \(N_j \subseteq N_{j+1}\), then there exists an index \(j_0 \in \N\)
  such that \(N_j = N_{j+1}\) for all \(j \geq j_0\).

\item Every non-empty collection of submodules of \(M\) has a \emph{maximal}
  element with respect to inclusion.
\end{enumerate}
\end{proposition}

\begin{proof}
\begin{itemize}\setlength\itemsep{0em}
\item (a) \(\implies\) (b). Let \(M\) be Noetherian and define the module
  \(N \coloneq \bigcup_{j \in \N} N_j\), which is a submodule of \(M\). Since
  submodules of Noetherian rings are finitely generated, it follows that \(N\)
  is finitely generated. Let \(N = \langle n_1, \dots, n_k \rangle\) be its
  generating set. For all \(1 \leq i \leq k\), there must exist \(j_i \in \N\)
  such that \(n_i \in N_j\) for all \(j \geq j_i\). Taking the maximum
  \(j_0 \coloneq \max(j_1, \dots, j_k)\), one finds that \(n_i \in N_j\) for
  each \(1 \leq i \leq k\) and every \(j \geq j_0\). Therefore
  \(N \subseteq N_j\) for all \(j \geq j_0\), which implies that \(N_j = N\) for
  each of those indexes --- therefore the chain stabilizes.

\item (b) \(\implies\) (c). We prove the contrapositive. Suppose there exists a
  non-empty collection \(\mathcal{N}\) of submodules of \(M\) admitting no
  maximal element. Let \(N_0 \in \mathcal{N}\) be any element. Inductively, for
  all \(j \geq 1\), define \(N_j \in \mathcal{N}\) such that \(N_{j-1}
  \subsetneq N_j\), that is, \(N_{j-1}\) is a \emph{proper} subset of \(N_j\)
  --- this is possible since \(N_{j-1}\) isn't maximal. The collection
  \((N_j)_{j \in \N}\) forms an ascending chain of submodules, but by
  construction does not stabilize.

\item (c) \(\implies\) (a). Let \(N \subseteq M\) be any submodule. Since
  \((0) \subseteq N\) is a finitely generated submodule of \(N\), one can define
  a non-empty collection \(\mathcal{N}\) of finitely generated submodules of
  \(N\). From (c) one has that \(\mathcal{N}\) admits a maximum element, say
  \(W \coloneq \langle n_1, \dots, n_k \rangle\). Let \(n \in N\) be any element
  and consider the finitely generated submodule \(\langle n_1, \dots, n_k, n
  \rangle \in \mathcal{N}\). Since \(W\) is maximal, we have \(\langle n_1,
  \dots, n_k, n \rangle \subseteq W\) --- therefore \(n \in W\) and \(N
  \subseteq W\). Therefore \(N = W\) is finitely generated, which proves that
  \(M\) is Noetherian.
\end{itemize}
\end{proof}

\begin{corollary}
\label{cor:pid-is-noetherian}
Every principal ideal domain \(R\) is a Noetherian module over itself, and thus
every non-empty collection of ideals of \(R\) admits a maximal element.
\end{corollary}

\begin{lemma}[Quotient of Noetherian rings is Noetherian]
\label{lem:quotient-is-noetherian}
Let \(R\) be a Noetherian ring, and \(\ideal{a} \subseteq R\) be an ideal. Then
the quotient ring \(R/\ideal{a}\) is Noetherian.
\end{lemma}

\begin{proof}
From \cref{prop:image-of-noetherian-is-noetherian} we find that the canonical
projection \(R \epi R/\ideal{a}\) implies that \(R/\ideal{a}\) is Noetherian.
\end{proof}

\begin{theorem}[Generalized Hilbert's basis theorem]
\label{thm:general-hilbert-basis}
Let \(R\) be a ring. Then \(R\) is Noetherian if and only if the polynomial ring
\(R[x_1, \dots, x_n]\).
\end{theorem}

\todo[inline]{Prove generalized Hilbert's basis theorem}

\begin{corollary}
\label{cor:noetherian-quotient-poly-ring}
Let \(R\) be a Noetherian ring, and \(\ideal{a} \subseteq R[x_1, \dots, x_n]\)
be an ideal of the polynomial ring. Then the quotient ring
\(R[x_1, \dots, x_n]/\ideal{a}\) is Noetherian.
\end{corollary}

\begin{proof}
Since \(R[x_1, \dots, x_n]\) is Noetherian by \cref{thm:general-hilbert-basis},
applying \cref{lem:quotient-is-noetherian} we find that \(R[x_1, \dots,
x_n]/\ideal a\) is Noetherian.
\end{proof}

\begin{corollary}
\label{cor:finite-type-alg-noetherian}
Every \emph{finite-type} algebra over a Noetherian ring is \emph{Noetherian}.
\end{corollary}

\subsection{Existence of Maximal Ideals}

\begin{proposition}
\label{prop:commutative-rings-have-maximal-ideals}
Let \(R\) be a commutative ring. If \(\ideal a\) is any \emph{proper} ideal of
\(R\), then there exists a \emph{maximal ideal} \(\ideal m\) of \(R\) containing
\(\ideal a\).
\end{proposition}

\begin{proof}
Consider the collection \(I\) of proper ideals of \(R\) containing
\(\ideal a\)---which is ordered by inclusion. Using this ordering, let
\((\ideal a_j)_{j \in J}\) be the chain of all proper ideals with
\(\ideal a_j \in I\) and \(\ideal a_j \subseteq \ideal a_{j+1}\). Define the set
\(\ideal m \coloneq \bigcup_{j \in J} \ideal a_j\), which is again an ideal of
\(R\). Since every \(\ideal a_j\) contains \(\ideal a\) and does \emph{not}
contain \(1\), it follows that \(\ideal m\) contains \(\ideal a\) and also
doesn't contain \(1\)---hence \(\ideal m\) is a proper ideal. Therefore
\(\ideal m\) is a \emph{maximal} ideal of \(R\) containing \(\ideal a\), which
proves the statement.
\end{proof}

\section{Localization}

\begin{definition}[Multiplicative subset]
\label{def:multiplicative-subset}
Given a commutative ring \(R\), a subset \(S \subseteq R\) is said to be
\emph{multiplicative} if \(1_R \in S\) and \(S\) is closed under
multiplication---that is, given \(s, s' \in S\), we have \(s s' \in S\).
\end{definition}

\begin{definition}[Localization]
\label{def:localization-ring}
Let \(R\) be a commutative ring, and \(S \subseteq R\) be a \emph{multiplicative
  subset}. A \emph{localization} of \(R\) over the set \(S\) is a morphism of
\emph{commutative rings} \(L_S: R \to R[S^{-1}]\) such that:
\begin{enumerate}[(a)]\setlength\itemsep{0em}
\item For all \(s \in S\), the image \(L_S(s) \in R[S^{-1}]\) is a \emph{unit}.
\item For every morphism of rings \(f: R \to K\) satisfying property (a), there
  exists a \emph{unique} ring morphism \(\phi: R[S^{-1}] \to K\) such that the
  diagram
  \[
  \begin{tikzcd}
  R \ar[r, "L_S"] \ar[rd, bend right, "f"']
  &{R[S^{-1}]} \ar[d, "\phi", dashed] \\
  &K
  \end{tikzcd}
  \]
\end{enumerate}
Equivalently, consider the minimal equivalence relation \(\sim_{\text{frac}}\)
on the set \(R \times S\) defined by \((r, s) \sim_{\text{frac}} (r', s')\) if
and only if \(r s' - s r' = 0\). Then the localization of \(R\) under the
multiplicative subset \(S\) is simply
\[
R[S^{-1}] = (R \times S)/{\sim_{\text{frac}}}.
\]
\end{definition}

\begin{example}
\label{exp:Z[1/p]/Z-finite-submodules}
Let \(p \in \Z\) be a prime number, and consider the ring of fractions
\(\Z[p^{-1}]\). Then every proper submodule of the \(\Z\)-module given by the
quotient \(\Z[p^{-1}]/\Z\) is \emph{finite}.
\todo[inline]{How to solve this?}
\end{example}

\begin{definition}
\label{def:field-of-fractions}
Let \(R\) be an integral domain. We define the \emph{field of fractions} of
\(R\) to be the localization
\[
\Frac(R) \coloneq R[(R \setminus \{0\})^{-1}].
\]
\end{definition}

It is immediate that \(\Frac(R)\) is a field, since a field is simply an
integral domain whose non-zero elements are units---and this exactly what we did
with the above localization.

\section{Principal Ideal Domains}

\subsection{Modules Over PID's}

\begin{theorem}
\label{thm:module-over-PID}
Let \(R\) be a principal ideal domain, and \(M\) be a free \(R\)-module with
\(\rank_R M = m\) finite, and \(N \subseteq M\) a submodule. Then:
\begin{enumerate}[(a)]\setlength\itemsep{0em}
\item The submodule \(N\) is free with \(\rank_R N = n\) satisfying \(n \leq m\).
\item There exists a basis \((y_1, \dots, y_m)\) of \(M\), and a collection
  \((a_1, \dots, a_m)\) of \(R\) such that \((a_1 y_1, \dots, a_n y_n)\) is a
  basis for \(N\) and \(a_1 \mid a_2 \mid \dots \mid a_n\).
\end{enumerate}
\end{theorem}

\begin{proof}
Fix a basis \((x_1, \dots, x_m)\) for \(M\) and define a collection of
\(R\)-linear morphisms \((\pi_j: M \epi R)_{j=1}^n\) with
\(\pi_j x_i = \delta_{i j}\).
\begin{enumerate}[(a)]\setlength\itemsep{0em}
\item Since \(R\) is a domain, \(M\) is torsion-free and so is \(N\). Since
  \(M\) has a finite rank, it must also be the case that \(N\) has a finite rank
  \(n\) satisfying \(n \leq m\)---which follows from
  \cref{prop:LI-leq-maximal-LI}. We proceed via induction on \(n\): for the base
  case \(n = 0\) we have \(N = 0\) and therefore \(N\) is free. We assume for
  the induction hypothesis that the proposition is true for some
  \(0 < n-1 < m\). Now we consider the case \(0 < n \leq m\): since \(n\) is
  non-zero, then \(N\) is a non-zero module and therefore there exists a
  non-zero \(x = \sum_{j=1}^n b_j x_j \neq 0\) element of \(N\). If
  \(1 \leq j_0 \leq n\) is an index such that \(b_{j_0} \neq 0\), then in
  particular \(\pi_{j_0}|_N\) is a non-zero map of the form \(N \to R\). Since
  \(R\) is a PID and \(\pi_{j_0} N \subseteq R\) is a non-zero submodule, then
  there must exist \(b_0 \in R \setminus 0\) for which \(\pi_{j_0} N = R
  b_0\). Since \(R\) is a domain, then \(\{b_0\}\) is a basis for the free
  \(R\)-module \(\pi_{j_0} N\). Then the short exact sequence
  \[
  \begin{tikzcd}
  0 \ar[r]
  &\ker(\pi_{j_0}|_N) \ar[r, hook]
  &N \ar[r, two heads, "\pi_{j_0}|_N"]
  &R b_0 \ar[r]
  &0
  \end{tikzcd}
  \]
  ends with a free module and thus splits, hence
  \[
  N \iso \ker(\pi_{j_0}|_N) \oplus R b_0.
  \]
  Since \(\rank\) is additive over direct sums, then
  \(n = \rank_R(\ker(\pi_{j_0}|_N)) + 1\) implies that
  \(\ker(\pi_{j_0}|_N) \subseteq M\) is a submodule of rank \(n-1\). From the
  inductive hypothesis it follows that \(\ker(\pi_{j_0}|_N)\) is a free
  module. Therefore \(N\) is the direct sum of two free modules, hence \(N\) is
  itself free.

\item Notice that the case \(M = 0\) is trivial, thus we shall consider only
  \(m \geq 1\) and do induction on \(m\)---moreover, we'll assume that \(N\) is
  anon-zero submodule of \(M\), since the zero case is trivially satisfied. For
  the base case \(m = 1\), there exists an \(R\)-module isomorphism
  \(\phi: M \isoto R\). Then from the fact that \(R\) is a PID it follows that
  there exists \(a \in R\) such that \(\phi N = R a = a R\), and if
  \(u \coloneq \phi^{-1}(1)\) then \(M = R u\)---so that \(\{u\}\) is a basis
  for \(M\). Moreover,
  \[
  N = a (R u) = (a R) u = (R a) u = R (a u),
  \]
  therefore \(\{a u\}\) forms a basis for \(N\).

  Now let \(m > 1\) and assume as a hypothesis for induction that the
  proposition is true for all modules of rank less than \(m\). For each
  \(\phi \in \Hom_{\Mod R}(M, R)\), let \(a_{\phi} \in R\) denote a ring element
  such that \(\phi N = R a_{\phi}\), and define a collection of principal ideals
  \[
  \Sigma \coloneq (R a_{\phi})_{\phi \in \Hom(M, R)}.
  \]
  From our previous considerations in item (a), we know that there must exist a
  non-zero \(\pi_{j_0} \in \Hom(M, R)\)---thus \(\Sigma\) is non-empty and its
  elements are not all zero. Since \(R\) is Noetherian, it follows that
  \(\Sigma\) admits a maximal element \(R a_{\phi_1} \in \Sigma\). Define
  \(a_1 \coloneq a_{\phi_1}\) and let \(y \in N\) be such that
  \(\phi_1 y = a_1\), which exists because \(\phi_1 N = R a_1\). We'll show that
  \(a_1 \mid \varphi y\) for any \(\varphi \in \Hom(M, R)\). Let
  \(\varphi: M \to R\) be any such map, and let
  \(\ideal p \coloneq \langle a_1, \varphi y \rangle\) be an ideal of \(R\), so
  that there exists \(d \in R\) such that \(\ideal p = R d\) since \(\ideal p\)
  is principal. Therefore there exists a pair \(r_1, r_2 \in R\) such that
  \(d = r_1 a_1 + r_2 \varphi y\). Define a morphism \(\psi: M \to R\) given by
  \(\psi \coloneq r_1 \phi_1 + r_2 \varphi\) so that \(\psi y = d\) and hence
  \(d \in \psi N\) from the fact that \(y \in N\). Since \(a_1 \in R d\), then
  \(R a_1 \subseteq R d\) and from maximality we obtain the equality
  \(R a_1 = R d\), which proves that \(a_1 \mid d\)---while
  \(d \mid \varphi y\), thus \(a_1 \mid \varphi y\) as we wanted.

  In particular, the last paragraph shows that \(a_1 \mid \pi_j y\) for each
  projection \(1 \leq j \leq m\), thus we may find \(b_j \in R\) such that
  \(\pi_j y = a_1 b_j\). We shall define
  \[
  y_1 \coloneq \sum_{j=1}^m b_j x_j.
  \]
  Notice then that \(y\) can be rewritten as
  \begin{equation}\label{eq:y=a1y1-rewrite}
  y = \sum_{j=1}^m \pi_j(y) x_j
  = \sum_{j=1}^m (a_1 b_j) x_j
  = a_1 \sum_{j=1}^m b_j x_j
  = a_1 y_1.
  \end{equation}
  Furthermore, one knows that \(a_1 = \psi y = \psi (a_1 y_1) = a_1 \psi y_1\),
  therefore since \(R\) is a domain it follows that \(\psi y_1 = 1\).

  We'll now show that \(M = R y_1 \oplus \ker \psi\). Let \(x \in M\) be any
  element and write \(x = \psi(x) y_1 + (x - \psi(x) y_1)\). Notice that
  \begin{equation}\label{eq:x-psixy1-in-kernel}
  \psi(x - \psi(x) y_1) = \psi x - \psi(x) \psi(y_1) = 0
  \end{equation}
  so \(x - \psi(x) y_1 \in \ker \psi\). Therefore \(M = R y_1 + \ker \psi\). Now
  if \(r y_1 \in \ker \psi\) then from the fact that \(\psi y_1 = 1\) we have
  \(r = 0\), hence \(r y_1 = 0\) and thus \(R y_1 \cap \ker \psi = 0\). We thus
  conclude that \(M = R y_1 \oplus \ker \psi\).

  Further, we'll prove that \(N = R (a_1 y_1) \oplus (N \cap \ker \psi)\). Let
  \(x \in N\) be any element and as before rewrite it as \(x = \psi(x) y_1 + (x
  - \psi(x) y_1)\). Using the fact that \(\psi N = R a_1\), there exists \(b \in
  R\) such that \(\psi x = b a_1\)---then one has
  \[
  x = (b a_1) y_1 + (x - (b a_1) y_1)
  \]
  where \(b a_1 y_1 \in R (a_1 y_1)\) and \(x - (b a_1) y_1 \in N \cap \ker \psi\)
  from \cref{eq:x-psixy1-in-kernel}. Therefore \(N = R (a_1 y_1) + (N \cap \ker
  \psi)\), and since \(R(a_1 y_1) \subseteq R y_1\) while \(N \cap \ker \psi
  \subseteq \ker \psi\) then \(R(a_1 y_1) \cap (N \cap \ker \psi) = 0\). Thus
  \[
  N = R(a_1 y_1) \oplus (N \cap \ker \psi).
  \]

  Since \(\ker \psi \subseteq M\) then it is free and the rank of the direct sum
  yields \(\rank_R \ker \psi = m - 1\). Thus from the inductive hypothesis we
  conclude that there exists a basis \((y_2, \dots, y_m)\) of \(\ker \psi\) and
  a collection of ring elements \((a_2, \dots, a_n)\) such that
  \((a_2 y_2, \dots, a_n y_n)\) forms a basis for \(N \cap \ker \psi\), and
  \(a_2 \mid a_3 \mid \dots \mid a_m\). Since \(M = R y_1 \oplus \ker \psi\)
  then \((y_1, y_2, \dots, y_m)\) is a basis for \(M\), and since \(N = R (a_1
  y_1) \oplus (N \cap \ker \psi)\) we have that \((a_1 y_1, a_2 y_2, \dots, a_n
  y_n)\) is a basis for \(N\).

  To conclude the proof, we must show that \(a_1 \mid a_2\). Since
  \((y_1, \dots, y_m)\) is a basis of \(M\), let \(f: M \to R\) be an
  \(R\)-linear morphism such that \(f y_1 = f y_2 = 1\) while \(f y_j = 0\) for
  each \(3 \leq j \leq m\). From \cref{eq:y=a1y1-rewrite} we have
  \(a_1 = a_1 f y_1 = f(a_1 y_1) = f y\), therefore since \(y \in N\) then \(a_1
  \in f N\). This implies in \(R a_1 \subseteq f N\) but using maximality this
  yields the equality \(f N = R a_1\). We also know that \(a_2 = f(a_2 y_2)\) in
  a similar fashion, and since \(a_2 y_2 \in N\) then \(a_2 \in f N\). This
  concludes the proof that \(a_1 \mid a_2\), which was the last step for the end
  of the proof at large.
\end{enumerate}
\end{proof}

\subsection{Decomposition Theorems}

\begin{theorem}[Fundamental theorem of invariant factors]
\label{thm:fundamental-theorem-of-invariant-factors}
Let \(R\) be a principal ideal domain and \(M\) a finitely generated
\(R\)-module. Then the following holds:
\begin{enumerate}[(a)]\setlength\itemsep{0em}
\item There exists a number \(r \in \N\) and a collection of non-zero and
  non-invertible elements \((a_1, \dots, a_m)\) of \(R\) such that
  \[
  M \iso R^r \oplus \Big( \bigoplus_{j=1}^m R/R a_j \Big),
  \]
  and \(a_1 \mid a_2 \mid \dots \mid a_m\). That is, \(M\) is isomorphic to the
  direct sum of a finite collection of cyclic \(R\)-modules. The elements
  \(a_j\) are said to be invariant factors of \(M\).

\item The module \(M\) is torsion-free if and only if \(M\) is free.

\item There exists an isomorphism
  \[
  \torsion M \iso \bigoplus_{j=1}^m R/Ra_j.
  \]
  In particular, \(M\) is a torsion module if and only if \(r = 0\), and in such
  a case we have \(\Ann M = R a_m\)\footnote{The annihilator of a module \(M\)
    is defined to be the collection
    \[
    \Ann M \coloneq \{r \in R \colon r m = 0 \text{ for all } m \in M\}.
    \]}.
\end{enumerate}
\end{theorem}

\begin{proof}
Let \(M = \langle x_1, \dots, x_n \rangle\) and let
\(e_j \coloneq (\delta_{ij})_{i=1}^n \in R^n\) be a base element of \(R^n\). Define a morphism
\(\phi: R^n \to M\) as the map that sends \(e_j \mapsto x_j\) for each
\(j\)---which is unique by the free module universal property. Since
\(\{x_1, \dots, x_n\}\) generates \(M\), then \(\im \phi = M\) and \(\phi\) is thus
surjective. Via the first isomorphism theorem we obtain
\(R^n/\ker \phi \iso M\). Since \(\ker \phi\) is a submodule of \(R^n\) we can use
\cref{thm:module-over-PID} to obtain that \(\ker \phi\) is free and has
\(\rank(\ker \phi) \coloneq m \leq n\). Furthermore, there exists a basis
\((t_1, \dots, t_n)\) of \(R^n\) and a list of ring elements
\((a_1, \dots, a_m)\) such that \((a_1 t_1, \dots, a_m t_m)\) is a basis of
\(\ker \phi\) satisfying \(a_1 \mid a_2 \mid \dots \mid a_m\). From the decomposition of
free modules we obtain
\[
M \iso R^n/\ker \phi =
\frac{R t_1 \oplus \dots \oplus R t_n}{R a_1 t_1 \oplus \dots \oplus R a_m t_m}
\]
Define a morphism of \(R\)-modules
\(\psi: \bigoplus_{j=1}^n R t_j \to \big( \bigoplus_{j=1}^m R/Ra_j \big) \oplus R^{n-m}\)
by mapping
\[
\psi(r_1 t_1, \dots, r_n t_n) \coloneq
(r_1 + R a_1, \dots, r_m + R a_m, r_{m+1}, \dots, r_n).
\]
From the definition, this map is certainly surjective. Moreover, we can see that
\[
\ker \psi
= \Big( \bigoplus_{j=1}^m R a_j \Big) \oplus \{0\}^{\oplus (n-m)}
\iso \bigoplus_{j=1}^m R a_j,
\]
therefore by the first isomorphism theorem we find
\[
\frac{R t_1 \oplus \dots \oplus R t_n}{R a_1 t_1 \oplus \dots \oplus R a_m t_m}
\iso \Big( \bigoplus_{j=1}^m R/R a_j \Big) \oplus R^{n-m}.
\]
Notice that if \(a \in R\) is an invertible element, then \(R a = R\) and hence
\(R/Ra = 0\). Therefore the proposition of item (a) is stablished since we can
cut off of the direct sum every summand \(R/Ra_j\) where \(a_j\) is invertible.
Let \(b \in R\) be any element, then given any \(r + R b \in R/Rb\) we find that
\[
b(r + R b) = b r + R b = r b + R b = R b
\]
therefore \(R/Rb\) is a torsion \(R\)-module. This shows that \(M\) is
torsion-free if and only if \(M \iso R^r\), which proves item (b). As we just
noted, one has \(\torsion M \iso \bigoplus_{j=1}^m R/R a_j\), and \(M\) is a
torsion module we have \(\Ann\)
\todo[inline]{Notice that this isn't finished: it is part of the lost
  work---this chapter was the most affected by the losses.}
\end{proof}

\subsection{Chinese Remainder Theorem}

\begin{theorem}[Chinese remainder]
\label{thm:chinese-remainder-theorem}
Let \(R\) be a commutative ring, and \(\ideal{a}_1, \dots, \ideal{a}_k\)
be ideals of \(R\) such that \(\ideal{a}_i + \ideal{a}_j = R\) for all
\(i \neq j\). Then:
\begin{itemize}\setlength\itemsep{0em}
\item We have the equality \(\ideal{a}_1 \cap \dots \cap \ideal{a}_k =
  \ideal{a}_1 \cdot \ldots \cdot \ideal{a}_k\).
\item The natural projection
  \(R \epi R/\ideal{a}_1 \times \dots \times R/\ideal{a}_k\) is
  \emph{surjective}, and induces a natural \emph{isomorphism} of rings
  \[
  \frac{R}{\ideal{a}_1 \dots \ideal{a}_k}
  \iso (R/\ideal{a}_1) \times \dots \times (R/\ideal{a}_{k})
  \]
\end{itemize}
\end{theorem}

\begin{proof}
In particular, we have \(\ideal{a}_j + \ideal{a}_k = R\) for all \(1 \leq j \leq
k - 1\). Therefore, for all such indices there exists \(a_j \in \ideal{a}_k\)
such that \(1 - a_j \in \ideal a_j\), hence
\[
(1 - a_1) \dots (1 - a_{k-1}) \in
\ideal{a}_1 \cdot \ldots \cdot \ideal a_{k-1},
\]
and since \(a_j \in \ideal{a}_k\), then
\(1 - \prod_{j=1}^{k-1} (1 - a_j) \in \ideal{a}_k\). This shows that
\begin{equation}\label{eq:coprime-ideals-CRT}
(\ideal{a}_1 \cdot \ldots \cdot \ideal{a}_{k-1}) + \ideal{a}_k = R.
\end{equation}

Notice that
\(\ideal{a}_1 \cdot \ldots \cdot \ideal a_k \subseteq \ideal a_1 \cap \dots \cap
\ideal a_k\), thus it remains to prove the other side of the inclusion. From our
last paragraph, we know that
\(\ideal a_1 \cap \dots \cap \ideal a_k \subseteq \ideal{a}_1 \cdot \ldots \cdot
\ideal a_k \) for all \(k \geq 3\). Since \(k = 1\) is trivial, we just need to
prove the case for \(k = 2\). Let \(\ideal{b}, \ideal{c} \subseteq R\) be ideals
such that \(\ideal b + \ideal c = R\)---then there exists \(b_0 \in \ideal b\)
and \(c_0 \in \ideal c\) such that \(b_0 + c_0 = 1\). If \(x \in \ideal{b} \cap
\ideal c\), then \(x = b_0 x + c_0 x\), implying in \(x \in \ideal b \cdot
\ideal c\). Thus \(\ideal b \cap \ideal c \subseteq \ideal b \cdot \ideal c\).

We now prove the second assertion via induction on \(k\). For the base case
\(k = 1\), the statement follows trivially from the first isomorphism
theorem. Assume as the hypothesis of induction that the statement is true
\(k - 1 > 1\)---that is, we have an isomorphism
\(R/(\ideal a_1 \cdot \ldots \cdot \ideal a_{k-1}) \iso (R/\ideal a_1) \times
\dots \times (R/\ideal a_{k-1})\). Consider the natural map
\begin{equation}\label{eq:natural-projection-CRT}
\pi: R \longrightarrow \frac{R}{\ideal a_{1} \cdot \ldots \cdot \ideal a_{k-1}}
\times (R/\ideal a_k).
\end{equation}
From \cref{eq:coprime-ideals-CRT} the statement is reduced for the case of two
ideals \(\ideal b \coloneq \ideal a_1 \cdot \ldots \cdot a_{k-1}\) and
\(\ideal c \coloneq \ideal a_k\) such that \(\ideal b + \ideal c = R\). Let
\(b, c \in R\) be any two elements---we shall show that there exists \(r \in R\)
such that \(r - b \in \ideal b\) and \(r - c \in \ideal c\). Since \(\ideal b\)
and \(\ideal c\) are relatively prime, let as before \(b_0 + c_0 = 1\) and
define \(r \coloneq b_0 c + c_0 b\). Then, one has
\begin{align*}
  r &= b_0 c + (1 - b_0) b = b + b_0 (c - b) \equiv b \pmod{\ideal b} \\
  r &= (1 - c_0) c + c_0 b = c + c_0 (b - c) \equiv c \pmod{\ideal c}
\end{align*}
since \(b_0 \in \ideal b\) and \(c_0 \in \ideal c\). Therefore
\(r - c \in \ideal b\) and \(r - b \in \ideal c\) a wanted. This shows that the
natural morphism of rings \(\pi\) (see \cref{eq:natural-projection-CRT}) is
surjective. From our first considerations, we know that
\(\ker \pi = \ideal b \cap \ideal c = \ideal b \cdot \ideal c\), therefore the
first isomorphism theorem establishes that
\(R/(\ideal b \cdot \ideal c) \iso (R/\ideal b) \times (R/\ideal c)\).
\end{proof}

\begin{corollary}[Chinese remainder for PIDs]
\label{cor:chinese-remainder-for-PID}
Let \(R\) be a principal ideal domain, and \(a_1, \dots, a_k \in R\) be elements
such that \(\gcd(a_i, a_j) = 1\) for all pairs \(i \neq j\). Then the natural
map \(r + (a_1 \cdots a_k) \mapsto (r + (a_1), \dots, r + (a_k))\) establishes an
\emph{isomorphism} of rings
\[
R/(a_1 \cdots a_k) \iso \frac{R}{(a_{1})} \times \dots \times \frac{R}{(a_k)}.
\]
\end{corollary}

\section{Unique Factorisation Domains}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../deep-dive"
%%% End:
